---
title: "Happiness"
author: "Shuyi Bian, Carmina Inguito, Yutong Ji"
date: "`2022-03-03`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction 

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(leaflet)
library(rgdal)
library(ggplot2)
library(maps)
library(rworldmap)
library(ggmap)
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(ggthemes)
library(shiny)
library(shinydashboard)
library(ggVennDiagram)
```

```{r}
options(scipen=999)
```


## Laod and Clean the Data
```{r}
happiness_2015 <- readr::read_csv("data/2015.csv")%>% 
  clean_names()
happiness_2016 <- readr::read_csv("data/2016.csv")%>% 
  clean_names()
happiness_2017 <- readr::read_csv("data/2017.csv")%>% 
  clean_names()
happiness_2018 <- readr::read_csv("data/2018.csv")%>% 
  clean_names()
happiness_2019 <- readr::read_csv("data/2019.csv")%>% 
  clean_names()
```
## Merging Data Frames
**Add year**
```{r}
# Add year
happiness_2015_join <- happiness_2015 %>% 
  mutate(year = case_when(country != "NA" ~ "2015"))
happiness_2016_join <- happiness_2016 %>% 
  mutate(year = case_when(country != "NA" ~ "2016"))
happiness_2017_join <- happiness_2017 %>% 
  mutate(year = case_when(country != "NA" ~ "2017"))
happiness_2018_join <- happiness_2018 %>% 
  mutate(year = case_when(country_or_region != "NA" ~ "2018"))
happiness_2019_join <- happiness_2019 %>% 
  mutate(year = case_when(country_or_region != "NA" ~ "2019"))
```

**Change column names**
```{r}
#Change Column names
happiness_2015_join <- happiness_2015_join %>% 
  rename(country_or_region = country)

happiness_2016_join <- happiness_2016_join %>% 
  rename(country_or_region = country)

happiness_2017_join <- happiness_2017_join %>% 
  rename(country_or_region = country)

happiness_2018_join <- happiness_2018_join %>% 
  rename(happiness_rank = overall_rank,
         happiness_score = score,
         economy_gdp_per_capita = gdp_per_capita,
         family = social_support,
         health_life_expectancy = healthy_life_expectancy,
         freedom = freedom_to_make_life_choices,
         trust_government_corruption = perceptions_of_corruption)

happiness_2019_join <- happiness_2019_join %>% 
  rename(happiness_rank = overall_rank,
         happiness_score = score,
         economy_gdp_per_capita = gdp_per_capita,
         family = social_support,
         health_life_expectancy = healthy_life_expectancy,
         freedom = freedom_to_make_life_choices,
         trust_government_corruption = perceptions_of_corruption)
```

**Merge data frames together**
```{r}
#merge data frames together
#thank you Joel
happiness_2018_join$trust_government_corruption <- as.numeric(happiness_2018_join$trust_government_corruption)

complete_happiness <- bind_rows(happiness_2015_join, happiness_2016_join, happiness_2017_join, happiness_2018_join, happiness_2019_join) #bind the data frames

complete_happiness <- complete_happiness %>% 
  select(-region, -standard_error, -dystopia_residual, -lower_confidence_interval, -upper_confidence_interval, -whisker_high, -whisker_low) #remove unwanted columns
```
## Exploring the Data 
**Here, we're interested in understanding what kind of data we're going to be working with by utilizing several functions to observer its structure.**
```{r}
summary(complete_happiness)
```

```{r}
glimpse(complete_happiness)
```

```{r}
complete_happiness %>%
  naniar::miss_var_summary()
```


##Intereactive Map of Global Happiness Score
```{r}
world_map <- readOGR( 
 dsn= paste0("data/TM_WORLD_BORDERS_SIMPL-0.3.shp"),
 layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)
```

```{r}
happiness <- happiness_2019_join %>%
  select(country_or_region, happiness_score, economy_gdp_per_capita, family, health_life_expectancy, freedom, generosity, trust_government_corruption)%>%
  rename(NAME="country_or_region")
happiness
```
**HELP HELP HELP HELP HELP**
```{r}
happiness2 <- happiness
renamed <- happiness2$NAME
renamed <- str_replace(renamed, c('Libya'), replacement = 'Libyan Arab Jamahiriya')
renamed <- str_replace(renamed, c('Venezuela Bolivarian Republic of'), replacement = 'Venezuela')
renamed <- str_replace(renamed, c('South Sudan'), replacement = 'Sudan')
renamed <- str_replace(renamed, c('Congo (Brazzaville)'), replacement = 'Congo')
renamed <- str_replace(renamed, c('Congo (Kinshasa)'), replacement = 'Democratic Republic of the Congo')
renamed <- str_replace(renamed, c('Tanzania'), replacement = 'United Republic of Tanzania')
renamed <- str_replace(renamed, c('Syria'), replacement = 'Syrian Arab Republic')
renamed <- str_replace(renamed, c('Iran'), replacement = 'Iran (Islamic Republic of)')
renamed <- str_replace(renamed, c('Moldova'), replacement = 'Republic of Moldavo')
renamed <- str_replace(renamed, c('North Moldavo'), replacement = 'Republic of Moldavo')
renamed <- str_replace(renamed, c('Laos'), replacement = 'Lao People Democratic Republic')
happiness2$NAME = renamed
```

```{r}
world_map@data <- left_join(world_map@data, happiness2, by="NAME")
```

```{r}
world_map@data$happiness_score[ which(world_map@data$happiness_score == 0)] = NA
world_map@data$happiness_score <- as.numeric(as.character(world_map@data$happiness_score))%>% round(2)
```

```{r}
mypalette <- colorNumeric( palette="viridis", domain=world_map@data$happiness_score, na.color="transparent")
mypalette(c(45,43))
```

```{r}
m <- leaflet(world_map)%>% addTiles()  %>% setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons( stroke = FALSE, fillOpacity = 0.5, smoothFactor = 0.5, color = ~colorNumeric("Set3", happiness_score)(happiness_score) )
```

```{r}
mybins <- c(0,2,4,6,8,10)
mypalette <- colorBin( palette="YlOrBr", domain=world_map@data$happiness_score, na.color="transparent", bins=mybins)
```

```{r}
mytext <- paste(
    "Country: ", world_map@data$NAME,"<br/>", 
    "GDP Contribution: ", world_map@data$economy_gdp_per_capita, "<br/>", 
    "Family Contribution: ", world_map@data$family, "<br/>",
    "Life Expectancy Contribution: ", world_map@data$health_life_expectancy, "<br/>",
    "Freedom Contribution: ", world_map@data$freedom, "<br/>",
    "Generosity Contribution: ", world_map@data$generosity, "<br/>",
    "Government Trust Contribution: ", world_map@data$trust_government_corruption, "<br/>",
    "Happiness Score: ", round(world_map@data$happiness_score, 2),
    sep="") %>%
  lapply(htmltools::HTML)
```

```{r}
leaflet(world_map) %>% 
  addTiles()  %>% 
  setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(happiness_score), 
    stroke=TRUE, 
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal=mypalette, values=~population, opacity=0.9, title = "Happiness Score", position = "bottomleft" )

m  
```

##Venn Diagram
```{r}
happiness_venn <- happiness_2019_join %>% 
  mutate(economy_contribution_percentage = economy_gdp_per_capita/happiness_score,
         family_contribution_percentage = family/happiness_score, 
         health_contribution_percentage = health_life_expectancy/happiness_score, 
         freedom_contribution_percentage = freedom/happiness_score, 
         generosity_contribution_percentage = generosity/happiness_score, 
         happiness_category = case_when(happiness_score <= 3 ~ "very unhappy",
                                                   happiness_score > 3 & happiness_score <= 5 ~ "unhappy",
                                                   happiness_score > 5 & happiness_score <= 7 ~ "happy",
                                                   happiness_score > 7 ~ "very happy"), 
         economy_category = case_when(economy_contribution_percentage < 0.1 ~ "insignificant",
                                      economy_contribution_percentage >= 0.1 ~ "significant"), 
         family_category = case_when(family_contribution_percentage < 0.1 ~ "insignificant",
                                      family_contribution_percentage >= 0.1 ~ "significant"),
         health_category = case_when(health_contribution_percentage < 0.1 ~ "insignificant",
                                      health_contribution_percentage >= 0.1 ~ "significant"),
         freedom_category = case_when(freedom_contribution_percentage < 0.1 ~ "insignificant", 
                                      freedom_contribution_percentage >= 0.1 ~ "significant"), 
         generosity_category = case_when(generosity_contribution_percentage < 0.1 ~ "insignificant",
                                      generosity_contribution_percentage >= 0.1 ~ "significant"))
```

```{r}
happy_country_vec <- happiness_venn %>%
  filter(happiness_category == "happy" | happiness_category == "very happy")%>%
  pull(country_or_region)
gdp_vec <- happiness_venn %>%
  filter(economy_category == "significant")%>%
  pull(country_or_region)
generosity_vec <- happiness_venn %>%
  filter(generosity_category == "significant")%>%
  pull(country_or_region)
family_vec <- happiness_venn %>%
  filter(family_category == "significant")%>%
  pull(country_or_region)
health_vec <- happiness_venn %>%
  filter(health_category == "significant")%>%
  pull(country_or_region)
freedom_vec <- happiness_venn %>%
  filter(freedom_category == "significant")%>%
  pull(country_or_region)
```

```{r}
why_happy_list <- list(happy_country_vec, gdp_vec, generosity_vec)
ggVennDiagram(why_happy_list, category.names = c("Country with Happy Score", "GDP Contirbution", "Generosity Contribution"))
```



