---
title: "Happiness"
author: "Shuyi Bian, Carmina Inguito, Yutong Ji"
date: "`2022-03-03`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction 

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(ggthemes)
library(shiny)
library(shinydashboard)
```

```{r}
options(scipen=999) #cancels the use of scientific notation for the session
```


## Laod and Clean the Data
```{r}
happiness_2015 <- readr::read_csv("data/2015.csv")%>% 
  clean_names()
happiness_2016 <- readr::read_csv("data/2016.csv")%>% 
  clean_names()
happiness_2017 <- readr::read_csv("data/2017.csv")%>% 
  clean_names()
happiness_2018 <- readr::read_csv("data/2018.csv")%>% 
  clean_names()
happiness_2019 <- readr::read_csv("data/2019.csv")%>% 
  clean_names()
```

## Exploring the Data 
**Here, we're interested in understanding what kind of data we're going to be working with by utilizing several functions to observer its structure.**
```{r}
summary(happiness_2016)
```


## Merging Data Frames
**Add year**
```{r}
# Add year
happiness_2015_join <- happiness_2015 %>% 
  mutate(year = case_when(country != "NA" ~ "2015"))
happiness_2016_join <- happiness_2016 %>% 
  mutate(year = case_when(country != "NA" ~ "2016"))
happiness_2017_join <- happiness_2017 %>% 
  mutate(year = case_when(country != "NA" ~ "2017"))
happiness_2018_join <- happiness_2018 %>% 
  mutate(year = case_when(country_or_region != "NA" ~ "2018"))
happiness_2019_join <- happiness_2019 %>% 
  mutate(year = case_when(country_or_region != "NA" ~ "2019"))
```

**Change column names**
```{r}
#Change Column names
happiness_2015_join <- happiness_2015_join %>% 
  rename(country_or_region = country)

happiness_2016_join <- happiness_2016_join %>% 
  rename(country_or_region = country)

happiness_2017_join <- happiness_2017_join %>% 
  rename(country_or_region = country)

happiness_2018_join <- happiness_2018_join %>% 
  rename(happiness_rank = overall_rank,
         happiness_score = score,
         economy_gdp_per_capita = gdp_per_capita,
         family = social_support,
         health_life_expectancy = healthy_life_expectancy,
         freedom = freedom_to_make_life_choices,
         trust_government_corruption = perceptions_of_corruption)

happiness_2019_join <- happiness_2019_join %>% 
  rename(happiness_rank = overall_rank,
         happiness_score = score,
         economy_gdp_per_capita = gdp_per_capita,
         family = social_support,
         health_life_expectancy = healthy_life_expectancy,
         freedom = freedom_to_make_life_choices,
         trust_government_corruption = perceptions_of_corruption)
```

**Merge data frames together**
```{r}
#merge data frames together
#thank you Joel
happiness_2018_join$trust_government_corruption <- as.numeric(happiness_2018_join$trust_government_corruption)

complete_happiness <- bind_rows(happiness_2015_join, happiness_2016_join, happiness_2017_join, happiness_2018_join, happiness_2019_join) #bind the data frames

complete_happiness <- complete_happiness %>% 
  select(-region, -standard_error, -dystopia_residual, -lower_confidence_interval, -upper_confidence_interval, -whisker_high, -whisker_low) #remove unwanted columns
```

## Happiness Contributors' Change 2015 - 2019
```{r}
#happiness factors change from 2015-2019
library(shiny)

complete_happiness$year <- as.character(complete_happiness$year)

ui <- dashboardPage(
  dashboardHeader(title = "Happiness"),
  dashboardSidebar(disable = T),
  dashboardBody(
  fluidRow(
  box(title = "Plot Options", width = 3,
  selectInput("factor", "Select Happiness Factors", choices = c("economy_gdp_per_capita", "family", "health_life_expectancy", "freedom", "trust_government_corruption"), 
              selected = "economy_gdp_per_capita"),
  helpText("Factors' values reflect the extent to which a factor contributes to the calculation of the Happiness Score")),
  
  box(title = "Plot of Happiness Factors Data", width = 7,
  plotOutput("plot", width = "600px", height = "500px")))))

server <- function(input, output, session) { 
  output$plot <- renderPlot({
  complete_happiness %>%
      group_by(year) %>% 
      summarize(mean_factor_value = mean(get(input$factor), na.rm=TRUE)) %>% 
      ggplot(aes_string(x="year", y="mean_factor_value", fill="year"))+
      geom_col()+
      geom_text(aes(label = round(mean_factor_value, 2)),
                position = position_dodge(width = 0.9),
                vjust = -0.5,
                color = "black",
                size = 5)+
      theme_linedraw()+
      scale_fill_brewer(palette = "Pastel1")+
      labs(x=NULL,
           y="mean factor value")+
      theme(axis.text.x = element_text(size = 15),
            axis.text.y = element_text(size = 15),
            axis.title.y = element_text(size = 18))

  })
  session$onSessionEnded(stopApp)
  }

shinyApp(ui, server)
```

## Relationship between Happiness Contributor and Happiness Score 2015 - 2019
```{r}
#happiness factors change from 2015-2019
library(shiny)

complete_happiness$year <- as.numeric(complete_happiness$year)

ui <- dashboardPage(
  dashboardHeader(title = "Happiness"),
  dashboardSidebar(disable = T),
  dashboardBody(
  fluidRow(
  box(title = "Plot Options", width = 3,
  selectInput("factor", "Select Happiness Factors", choices = c("economy_gdp_per_capita", "family", "health_life_expectancy", "freedom", "trust_government_corruption"), 
              selected = "economy_gdp_per_capita"),
  sliderInput("year",
              "Year:",
              min = 2015,
              max = 2019,
              value = 2015,
              step = 1),
  
  helpText("Factors' values reflect the extent to which a factor contributes to the calculation of the Happiness Score")),
  
  box(title = "Plot of Happiness Factors Data", width = 7,
  plotOutput("plot", width = "600px", height = "500px")))))

server <- function(input, output, session) { 
  output$plot <- renderPlot({
    complete_happiness %>% 
      filter(year==input$year) %>% 
      ggplot(aes_string(x=input$factor, y="happiness_score"))+
      geom_point(size = 4, alpha = 0.5)+
      geom_smooth(method=lm, se=T)+
      labs(y="Happiness Score")+
      theme_linedraw()+
      scale_color_brewer(palette = "Pastel1")+
      labs(x="factor value",
           y="happiness score")+
      theme(axis.text.x = element_text(size = 15),
            axis.text.y = element_text(size = 15),
            axis.title.y = element_text(size = 18),
            axis.title.x = element_text(size = 18))
      
    
  })
  session$onSessionEnded(stopApp)
}

shinyApp(ui, server)
```


## Data Visualization
```{r}
names(complete_happiness)
```
##Anlysis
```{r}
getwd()
```


##Research Questions for Analysis 